---
title: ""
subtitle: ""
author: ""
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: false
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: www/web_report.css
    editor_options:
      chunk_output_type: console
---

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


# Homeownership data

## Set up

Load libraries

```{r setup}
options(scipen = 999)

library(tidyverse)
library(readxl)
library(urbnthemes)
library(tidycensus)

set_urbn_defaults("print")
```

Create directories for data.
```{r data}
if (!dir.exists("data")) {
  dir.create("data")
}

```

## Download and clean CPS data

### Download

We will use two tables for this:

1. Table 14: Quarterly Homeownership Rates for the U.S. and Regions: 1964 to Present

2. Table B-5. Margins of Error of Homeownership Rates for the United States


```{r download}
if (!file.exists("data/histtab14.xlsx")) {
  
  download.file(url = "https://www.census.gov/housing/hvs/data/histtab14.xlsx",
                destfile = "data/histtab14.xlsx",
                mode="wb")
  
}

if (!file.exists("data/B5.xlsx")) {
  
  download.file(url = "https://www.census.gov/housing/hvs/data/B5.xlsx",
                destfile = "data/B5.xlsx",
                mode="wb")
  
}


```

### Read and clean

#### Homeownership rate

First, read in and clean up the homeownership rate data.

```{r horead}

rate <- read_excel("data/histtab14.xlsx",
                   skip = 3)

# Get rid of rows of footnotes
rate <- rate %>% 
  head(-6)

```

The year and region variables share a column. Extract the year bu coercing a new year variable into an integer (this will drop the region names), and fill each region with the correct year.

```{r fill-year}

rate <- rate %>% 
  mutate(year = as.integer(`Year and Area`)) %>% 
  filter(!is.na(`Year and Area`)) %>% 
  fill(year, .direction = "down") %>% 
  filter(str_detect(`Year and Area`, "[:digit:]") == FALSE) 

```

There are three years that have revisions, so they are duplicated. Get only the revised estimates- these are listed after the unrevised estimates. Get the second set of estimates, and merge them back onto the master data. 

```{r ho-revise}

# Get only second set
rev <- rate %>% 
  group_by(year) %>% 
  filter(n() > 5) %>% 
  mutate(num = row_number()) %>% 
  filter(num %in% 6:10) %>% 
  select(-num)

rate <- rate %>% 
  filter(!year %in% unique(rev$year)) %>% 
  bind_rows(rev)

stopifnot(
  rate %>% group_by(year) %>% filter(n() > 5) %>% nrow() == 0
)

rm(rev)

```

Fix the region names, reshape, format date variable, convert homeownership rate to percent.

```{r ho-regnames}

names(rate) <- c("region", "Q1", "Q2", "Q3", "Q4", "year")

rate <- rate %>% 
  mutate(region = str_remove_all(region, "[:punct:]")) %>% 
  gather(key = "quarter", value = "horate", -region, -year) %>% 
  arrange(region, year, quarter) %>% 
  mutate(period = paste0(year, quarter),
         date = lubridate::yq(period),
         horate = horate / 100)

```

#### Margin of error

Repeat a similar process for the margin of error.

```{r read-error}

margin <- read_excel("data/B5.xlsx", skip = 7,
                     col_names = c("year", "Q1", "Q2", "Q3", "Q4")) %>% 
  head(-5)

```

Clean up year variable, reshape, format date variable. The margin of error represents the 90 percent confidence interval in percent- divide by 100.

```{r clean-error}

margin <- margin %>% 
  mutate(year = str_remove_all(year, "[:punct:]")) %>% 
  gather(key = "quarter", value = "moe", -year) %>% 
  mutate(period = paste0(year, quarter),
         date = lubridate::yq(period),
         year = as.integer(year),
         moe = moe / 100)
  

```

### Combine

Filter only national observations, merge with margin of error. Select only observations after 2000. Create variables that represent the bottom and top of the 90% interval.

Calculate two additional variables- the year-over-year change, and the quarter over quarter change.


```{r combine}

cps <- rate %>% 
  filter(region == "United States",
         !is.na(horate)) %>% 
  left_join(margin, by = c("year", "quarter", "period", "date")) %>% 
  filter(year >= 2000) %>% 
  mutate(horate_low = horate - moe,
         horate_high = horate + moe,
         horate_yoy = horate / lag(horate, 4) - 1,
         horate_mom = horate - lag(horate, 1))

write_csv(cps, "data/cps-horate_20002017.csv")

```


## Download and clean the ACS

We will use `library(tidycensus)` to get a time series of homeownership rates and margins of error. The census API is not available for prior years- hard code the other years from [American Fact Finder](https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_17_1YR_B25003&prodType=table)

```{r tidycensus}

acs <- map_df(2012:2018, ~ mutate(get_acs(geography = "US",
                                 year = .x,
                                 survey = "acs1",
                                 variable = "B25003_002",
                                 summary_var = "B25003_001"),
                                 year = .x))

# Hard code estimates from other years from fact finder
acs <- acs %>% 
  bind_rows(tribble(
  ~year, ~estimate, ~moe, ~summary_est, ~summary_moe,
  2005, 74318982, 293104, 111090617, 143575,
  2006, 75086485, 218471, 111617402, 145530,
  2007, 75515104, 227236, 112377977, 144356,
  2008, 75373053,	224087, 113101329, 146859,
  2009, 74843004, 217682, 113616229, 161397,
  2010, 74873372,	216091, 114567419, 163249,
  2011, 74264435,	230440, 114991725, 179541
))


acs <- acs %>% 
  mutate(date = as.Date(paste0(year, "-12-01")),
         horate = estimate / summary_est,
         horate_low = (estimate - moe) / summary_est,
         horate_high = (estimate + moe) / summary_est) %>% 
  fill(GEOID, NAME, variable) %>% 
  arrange(year)

write_csv(acs, "data/acs-horate_20052018.csv")
```

## Combine CPS and ACS

Make a joint data set for plotting.

```{r cps-acs}

jointdata <- acs %>% 
  select(year, date,
         horate, horate_low, horate_high) %>% 
  bind_rows(select(cps, year,
                        horate, horate_low, horate_high))

write_csv(jointdata, "data/cps-acs-horate.csv")

```

