---
title: ""
subtitle: ""
author: ""
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: false
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: www/web_report.css
    editor_options:
      chunk_output_type: console
---

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


# Homeownership data

## Set up

Load libraries

```{r setup}
options(scipen = 999)

library(tidyverse)
library(readxl)
library(urbnthemes)
library(tidycensus)

set_urbn_defaults("print")
```

Create directories for data.
```{r data}
if (!dir.exists("data")) {
  dir.create("data")
}

```

## Download and clean CPS data

### Download

We will use two tables for this:

1. Table 14: Quarterly Homeownership Rates for the U.S. and Regions: 1964 to Present

2. Table B-5. Margins of Error of Homeownership Rates for the United States


```{r download}
if (!file.exists("data/histtab16.xlsx")) {
  
  download.file(url = "https://www.census.gov/housing/hvs/data/histtab16.xlsx",
                destfile = "data/histtab16.xlsx",
                mode="wb")
  
}

if (!file.exists("data/B8.xlsx")) {
  
  download.file(url = "https://www.census.gov/housing/hvs/data/B8.xlsx",
                destfile = "data/B8.xlsx",
                mode="wb")
  
}


```

### Read and clean

#### Homeownership rate

First, read in and clean up the homeownership rate data.

```{r horead}

rate <- read_excel("data/histtab16.xlsx",
                   skip = 3)

# Get rid of rows of footnotes
rate <- rate %>% 
  head(-8)

```

The year and region variables share a column. Extract the year bu coercing a new year variable into an integer and fill each region with the correct year.

```{r fill-year}

rate1 <- rate %>% 
  rename(asian = ...6) %>% 
  mutate(year = as.integer(str_sub(`Year and         Quarter`, 1, 4))) %>% 
  filter(!is.na(`Year and         Quarter`)) %>% 
  fill(year, .direction = "down") %>% 
  filter(!is.na(`U.S.`)) %>% 
  mutate(quarter = paste0("Q", str_sub(`Year and         Quarter`, 1, 1)),
         period = paste0(year, quarter),
         date = lubridate::yq(period)) %>% 
  mutate(region = "United States") %>% 
  select(region, year, quarter, period, date,
         total = `U.S.`,
         white = `Non-Hispanic White Alone`,
         black = `Black Alone`,
         asian,
         all_other = `All Other Races`,
         hispanic = `Hispanic (of any race)`) %>% 
  gather(key = "race", value = "horate", 
         -c(region, year, quarter, period, date)) %>% 
  mutate(horate = as.numeric(horate) / 100)
  
  

stopifnot(rate1 %>% 
            group_by(year, race) %>% 
            count() %>% 
            filter(n > 5) %>% 
            nrow() == 0)

stopifnot(rate1 %>% 
            group_by(year, race) %>% 
            count() %>% 
            filter(n < 4) %>% 
            nrow() != 1)

```



#### Margin of error

Repeat a similar process for the margin of error.

```{r read-error}

margin <- read_excel("data/B8.xlsx", skip = 5) %>% 
  head(-11)

names(margin) <- c("quarter", "white", "black", "all_other", "asian", "hispanic")

```

Clean up year variable, reshape, format date variable. The margin of error represents the 90 percent confidence interval in percent- divide by 100.

```{r clean-error}

margin1 <- margin %>% 
  mutate(year = as.integer(quarter)) %>% 
  fill(year, .direction = "down") %>% 
  filter(!is.na(white)) %>% 
  mutate(quarter1 = str_remove_all(quarter, "[:punct:]"),
         quarter = case_when(quarter1 == "First Quarter" ~ "Q1",
                             quarter1 == "Second Quarter" ~ "Q2",
                             quarter1 == "Third Quarter" ~ "Q3",
                             quarter1 == "Fourth Quarter" ~ "Q4"),
         period = paste0(year, quarter),
         date = lubridate::yq(period),
         asian = ifelse(asian == "(X)",
                        NA,
                        asian),
         region = "United States") %>% 
  select(region, year, quarter, period, date,
         white, black, all_other, asian, hispanic) %>% 
  gather(key = "race", value = "moe",
         -c(region, year, quarter, period, date)) %>% 
  mutate(moe = as.numeric(moe) / 100)
  

```

### Combine

Filter only national observations, merge with margin of error. Select only observations after 2000. Create variables that represent the bottom and top of the 90% interval.

Calculate two additional variables- the year-over-year change, and the quarter over quarter change.


```{r combine}

cps_race <- rate1 %>% 
  left_join(margin1, by = c("year", "quarter", "period", "date", "race", "region")) %>% 
  filter(year >= 2010) %>% 
  mutate(horate_low = horate - moe,
         horate_high = horate + moe,
         horate_yoy = horate / lag(horate, 4) - 1,
         horate_mom = horate - lag(horate, 1))


```


## Download and clean the ACS

We will use `library(tidycensus)` to get a time series of homeownership rates and margins of error. The census API is not available for prior years- hard code the other years from [American Fact Finder](https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_17_1YR_B25003&prodType=table)

```{r tidycensus}

acs <- map_df(2012:2018, ~ mutate(get_acs(geography = "US",
                                 year = .x,
                                 survey = "acs1",
                                 variable = "B25003_002",
                                 summary_var = "B25003_001"),
                                 year = .x,
                                 race = "total")) %>% 
  rbind(map_df(2012:2018, ~ mutate(get_acs(geography = "US",
                                 year = .x,
                                 survey = "acs1",
                                 variable = "B25003B_002",
                                 summary_var = "B25003B_001"),
                                 year = .x,
                                 race = "black")))

# Hard code estimates from other years from fact finder
acs <- acs %>% 
  bind_rows(tribble(
  ~year, ~estimate, ~moe, ~summary_est, ~summary_moe, ~race,
  2010, 74873372,	216091, 114567419, 163249, "total",
  2011, 74264435,	230440, 114991725, 179541, "total"
)) %>% 
  bind_rows(tribble(
  ~year, ~estimate, ~moe, ~summary_est, ~summary_moe, ~race,
  2010, 6070627,	39146, 13738492, 41846, "black",
  2011, 5977100,	35414, 14051208, 33702, "black"
))
  


maxyear <- max(acs$year)

acs <- acs %>% 
  mutate(date = as.Date(paste0(year, "-12-01")),
         horate = estimate / summary_est,
         horate_low = (estimate - moe) / summary_est,
         horate_high = (estimate + moe) / summary_est) %>% 
  fill(GEOID, NAME, variable) %>% 
  arrange(year)

```

## Combine CPS and ACS

Make a joint data set for plotting.

```{r cps-acs}

jointdata <- acs %>% 
  mutate(survey = "ACS") %>% 
  select(year, date, survey, race,
         horate, horate_low, horate_high) %>% 
  bind_rows(select(cps_race, year, date, race,
                        horate, horate_low, horate_high) %>% 
              mutate(survey = "CPS")) %>% 
  filter(race == "black")

write_csv(jointdata, "data/cps-acs-horate_black.csv")

```

